---
title: "SPY ETF ML Trading Strategy Analysis"
author: "Algorithmic Trading System"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: true
    toc_float: true
    theme: flatly
  pdf_document: 
    toc: true
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

# Data Load
The numeric suffix for each data set represents 1. ATR multiplier 2.
Trade timeout duration (minutes). View data_mining.ipynb for how this is populated.

```{r}
data8_60 = read.csv("../csvs/train8_60.csv")
```

# Data exploration for 8_60

```{r}
summary(data8_60)
```

## View Summary and shape

### ATR Multiplier (Lambda)

#### Mean Price Calculation:

$$\mu_T = \frac{\sum_{i=T-k}^{T}(OHLC_i)}{k}$$

#### Modified ATR Formula:

$$ATR_T = \sqrt{\frac{\sum_{i=T-k}^{T}(OHLC_i - \mu_T)^2}{k}}$$ Controls
the sensitivity to market volatility:

-   **Purpose**: Scales profit/loss targets based on current market
    conditions

-   **Example**: If SPY ATR = 0.50 and $\lambda = 10$:

    -   Base volatility adjustment = 10 × 0.50 = \$5.00

    -   Sell if SPY is up 5\$ from entry price

### pl_value

It is determined by iterating through the dataframe and seeing if the
price reaches take profit or stop loss first. It assumes a 1000 share
position size.

#### Profit-to-Loss Ratio (Chi)

Asymmetric risk-reward ratio: - **Default**: 1.5 (profit targets 50%
wider than loss targets)

-   **Profit Target**: $Price_T + \lambda * \chi × ATR$

-   **Loss Target**: $Price_T - \lambda * ATR$

#### Timeout Period (t)

Maximum holding period before forced exit:

-   **Purpose**: Prevents indefinite position holding

-   **Logic**: If neither target hit within t periods, compare exit
    price to entry

-   **Classification**: Profit (1) if $Price_{T+t} > Price_T$, Loss (0)
    otherwise

### SMA

$$SMA_k = \frac{Open_T}{\frac{1}{k}\sum_{i=T-k}^{T-1}Close_i}$$

-   $T$ = current time period

-   $k$ = lookback period

-   Hypothetically, values \> 1.0 indicate price above historical
    average (bullish)

-   Values \< 1.0 indicate price below historical average (bearish)

#### Multi-Timeframe Analysis

The system calculates SMA for three periods:

-   **SMA_7**: Short-term momentum (1 week of 5-min bars)

-   **SMA_20**: Medium-term trend (1 month of daily closes)

-   **SMA_50**: Long-term trend (quarterly trend)

#### Signal Interpretation

-   **Trend Confirmation**: Multiple SMAs above 1.0 = strong uptrend

-   **Momentum Divergence**: SMA_7 \> SMA_20 \> SMA_50 = accelerating
    uptrend

-   **Mean Reversion**: Extreme SMA values (\>1.05 or \<0.95) suggest
    potential reversal

See data_mining.ipynb for information

### PL: Encodes 0 if `pl_value` \< 0

Split train-test data

```{r}
seed_num = 1
set.seed(seed_num) # reproducibility
data8_60$norm_volume = (data8_60$Volume - mean(data8_60$Volume))/sd(data8_60$Volume)
ind = sample(1:nrow(data8_60),0.75*nrow(data8_60))
train = data8_60[ind,]
test = data8_60[-ind,]
```

Investigate dependent variable

```{r}
hist(train$pl_value, breaks = 50, col = "skyblue", main = "Distribution of pl_value (P&L)", xlab = "pl_value")
```

## Assess correlation

```{r}
par(mfrow=c(2,2))
plot(train$SMA_k7/train$SMA_k20, train$pl_value, main="SMA_k7 / SMA_k20 vs P&L", xlab="SMA_k7/SMA_k20", ylab="pl_value")
plot(train$SMA_k20/train$SMA_k50, train$pl_value, main="SMA_k20 / SMA_k50 vs P&L", xlab="SMA_k20/SMA_k50", ylab="pl_value")
plot(train$ATR, train$pl_value, main="ATR vs P&L", xlab="ATR", ylab="pl_value")
plot(log(train$norm_volume), train$pl_value, main="Volume vs P&L", xlab="normalized(Volume)", ylab="pl_value")
par(mfrow=c(1,1))
```

ATR shows to have a weak correlation. I think the interaction of
predictors with each other will be more important. Let's see the interaction of volume and ATR
```{r}
library(ggplot2)
ggplot(train, aes(x=ATR, y=norm_volume, color=PL)) +
  geom_point(alpha=0.5) +
  scale_color_gradient2(mid="red", high="blue") +
  theme_minimal() +
  ggtitle("ATR vs Volume")
```
This shows us that low ATR and volume have a lower chance of being profitable. Though it tends to fluctuate quite a lot.

## Visualize predictor correlation matrix

```{r}
library(corrplot)

numeric_vars <- train[, c("Open", "Close", "High", "Low", "norm_volume", "ATR", "SMA_k7", "SMA_k20", "SMA_k50", "pl_value")]
cor_matrix <- cor(numeric_vars, use="complete.obs")

corrplot(cor_matrix, method = "color", type = "upper", tl.cex = 0.8)
```
We will have to take steps to ensure we address multicollinearity, one way of doing this is looking at the VIF (Variance Inflation Factor). VIF is a diagnostic tool used to detect multicollinearity, which occurs when independent variables in a regression model are highly correlated with each other. High VIF values indicate that a variable's variance is inflated due to multicollinearity. Values over 5 may have potential issues in fitting models.

```{r}
library(car)
vif_model <- lm(pl_value ~ ATR + SMA_k7 + SMA_k20+ SMA_k50+ norm_volume, data=train)
vif(vif_model)
```
This shows us that in order to reduce multicollinearity, we must omit `SMA_k20`

## Log regression

### Model Diagnostics and Theoretical Foundations

Logistic regression models the log-odds of a binary outcome as a linear
function of predictors:

$$\log\left(\frac{p}{1-p}\right) = \beta_0 + \beta_1X_1 + \beta_2X_2 + ... + \beta_kX_k$$

Where $p$ is the probability of a profitable trade. The coefficients
represent the change in log-odds associated with a one-unit increase in
the predictor, holding all other variables constant. We use all parameters and all interactions initially

```{r}
glm1 = glm(
  data = train,
  PL ~ (ATR + SMA_k7 + SMA_k50 + norm_volume)^4,
  family = "binomial"
)

```

### Assess
The full model will likely have only a handful of terms with significant probabilities because of multi-colinearity
```{r}
summary(glm1)
```


### Residuals analysis

Regular residual plots don't make sense in glm. Dunn and Gordon (2018)
introduce quantile residuals for discrete response variables. Their
primary benefits are they do not show weird patterns (due to variable's
discreteness). They are available in R via statmod::qresid().

```{r}
library(statmod)
plot(density(qresid(glm1)))
```

This residuals do appear normally distributed. Which means there is no
unexplained variance (non linearity or omitted variables) in the model.

### Variable Selection

The Akaike Information Criterion (AIC) balances model fit against
complexity:

$$AIC = -2\ln(L) + 2k$$

Where $L$ is the likelihood and $k$ is the number of parameters. Lower
AIC values indicate better models. The stepwise procedure systematically
adds or removes variables to minimize AIC, providing a theoretically
sound approach to model selection. Let's use forward AIC to trim some of the predictors. AIC is usually
computationally expensive, however I am not dealing with a lot of
predictors here.

```{r}
glm_step <- step(glm1, direction = "both", trace = 0)
summary(glm_step)
plot(density(qresid(glm_step)))
```


This is still a lot of predictors, let's try and do some subset
selection. This method uses cross validation and fits models with the least mean squared error. By splitting the data up we are able to understand what is actually needed by minimizing MSE.

```{r}
library(glmnet)
# Get interaction matrix
X <- model.matrix(PL ~ (ATR + SMA_k7 + SMA_k50 + norm_volume)^4, data = train)[, -1]
y <- train$PL
cvfit <- cv.glmnet(X, y, family = "binomial", alpha = 1)

# coefs with least mse
lasso_coef <- coef(cvfit, s = "lambda.min")
selected_vars <- rownames(lasso_coef)[lasso_coef[,1] != 0][-1]  # exclude intercept
selected_vars
```

These are a lot fewer than earlier, let's fit this conservative model as well.

```{r}
glm_formula <- as.formula(paste("PL ~", paste(selected_vars, collapse = " + ")))
glm_cons <- glm(glm_formula, data = train, family = "binomial")
summary(glm_cons)
```

### Assess performance

```{r}
pred_glm = ifelse(predict(glm_step,test,type="response")>0.5,1,0)
winrate = mean(pred_glm == test$PL)
print(winrate)

pred_glm_cons = ifelse(predict(glm_cons,test,type="response")>0.5,1,0)
winrate_cons = mean(pred_glm_cons == test$PL)
print(winrate_cons)

if (winrate_cons < winrate) {
  glm2 = glm_step
} else {
  glm2 = glm_cons
}

```

The accuracy is
good, but it needs to be contextualized. The worst, bare bones model
would trade on a coin flip. Let's simulate $nrow(test)$ coin flips and
compare it to a 55% win rate.

Simulate 1000 trading simulations with random buy indicators and see if
winrate if better than it. `p = 0.05`. This is to weed out strategies
that don't perform better than random chance.

```{r}
n <- nrow(test)

# Simulate coin flips (baseline)
baseline_accuracy = c()
for (i in 1:1000) {
  set.seed(i)
  coin_flips <- rbinom(n, size = 1, prob = 0.5)
  baseline_accuracy <- c(mean(coin_flips == test$PL), baseline_accuracy)
}
set.seed(seed_num) # reproducibility
cat("Coin toss simulations that performed better than strategy :",length(baseline_accuracy[baseline_accuracy>winrate]) / length(baseline_accuracy))


# Use normal approximation

cat("\nWinrate p value:",pnorm(winrate, mean(baseline_accuracy), sd(baseline_accuracy), lower.tail = F))
```

This is definitely below the p value.

### Buy threshold

We need to determine when to buy based on the threshold. This improves
the win rate at the cost of fewer trades. This will miss out on a lot of
winning trades too (False negatives). But to succeed at trading we need
to minimize false positives. Precision, or $\frac{TP}{FP+TP}$ is much
more important. "For estimating a binomial proportion, at least 10
successes and 10 failures is recommended for normal approximation
intervals." Reference: Agresti, A. (2013). Categorical Data Analysis
(3rd ed.)

```{r}
# Define thresholds
probs <- 10^seq(-2, -10, length.out = 9)

fit_mean <- mean(glm2$fitted.values)
fit_sd <- sd(glm2$fitted.values)

prob_with_best_winrate = 0.05 # init
best_winrate = 0
for (prob in probs) {
  threshold <- qnorm(1 - prob, mean = fit_mean, sd = fit_sd)
  pred_glm <- ifelse(predict(glm_cons, train, type = "response") > threshold, 1, 0)

  true_positive <- sum(pred_glm == 1 & train$PL == 1)
  false_positive <- sum(pred_glm == 1 & train$PL == 0)
  total_predicted_positive <- sum(pred_glm == 1)
  if ((true_positive + false_positive) < 30) {
    break # CLT
  }
  winrate <- if (total_predicted_positive > 0) {
    true_positive / total_predicted_positive
  } else {
    NA
  }
  if (winrate > best_winrate) {
    best_winrate = winrate
    prob_with_best_winrate = prob
  }

  cat(sprintf("Prob = %e | TP = %d | FP = %d | Winrate = %.3f\n", 
              prob, true_positive, false_positive, winrate))
}

best_threshold_glm = qnorm(1 - prob_with_best_winrate, mean = fit_mean, sd = fit_sd)
cat("Best winrate", best_winrate*100)
cat("\nProbability with highest precision", prob_with_best_winrate)
cat("\nThreshold", best_threshold_glm)


```

### Drawdown

`pl_value` is calculated with 1000 shares of SPY. In trading, using 2%
of your portfolio in a trade is recommended, hence, The initial capital
is set at $\frac{1}{0.02} * (1000 * Price)$.

```{r}
initial_capital = (1/(0.02))*(1000 * sample(train$Open, 1)) # random price
capital <- initial_capital
drawdown <- c(capital)
wins <- c()

for (i in 1:nrow(test)) {
  pred_X <- predict(glm2, test[i, ], type = "response")
  if (pred_X > best_threshold_glm) {
    capital <- capital + test$pl_value[i]
    drawdown <- c(drawdown, capital)
    wins <- c(wins, test$pl_value[i])
  }
}

ggplot(data = data.frame(trade = 1:length(drawdown), 
                        equity = drawdown), 
       aes(x = trade, y = equity)) +
  geom_line(color = "blue", size = 1) +
  theme_minimal() +
  labs(title = "Equity Curve",
       x = "Trade #",
       y = "Capital") +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey95")
  ) +
  scale_y_continuous(labels = scales::comma)

# Sharpe ratio (simplified version: mean return over SD of returns)
sharpe <- (tail(drawdown, 1) - initial_capital) / sd(wins)


peak = max(drawdown)
trough = min(drawdown)
max_drawdown = (trough-peak)/peak


cat("Initial Capital", initial_capital)
cat("\nFinal Capital:", tail(drawdown, 1), "\n")
cat("Sharpe Ratio:", round(sharpe, 3), "\n")
cat("Max Drawdown:", max_drawdown*100, "%\n")
cat("Returns:", ((capital/initial_capital)-1)*100, "%\n")
```

## XGboost

In this section, we use **XGBoost**, a gradient-boosted decision tree
ensemble, to classify trade direction. Our goal is not just accuracy but
**precision**, the ratio of true positives to all predicted positives.
This focus reduces false positives—critical for real-world trading where
bad trades are costly.

### Load Data

We construct a design matrix using all two- and three-way interactions
of trading indicators. The outcome variable `PL` is a binary indicator
for whether a trade was profitable.

```{r}
library(xgboost)
library(Matrix)
library(caret)

data = train
X = model.matrix(PL ~ (ATR + SMA_k7 + SMA_k50 + norm_volume)^4, data=data)[, -1]
y = data$PL
dtrain = xgb.DMatrix(data = X, label = y)
```

### XGBoost Theoretical Framework

XGBoost builds an ensemble of decision trees sequentially, with each new
tree correcting errors made by previous trees. The algorithm optimizes:

$$\mathcal{L}(\phi) = \sum_{i=1}^n l(y_i, \hat{y}_i) + \sum_{k=1}^K \Omega(f_k)$$

Where: - $l$ is a differentiable convex loss function - $\hat{y}_i$ is
the prediction - $\Omega$ is a regularization term controlling model
complexity - $f_k$ represents the $k$-th tree in the ensemble

This approach allows XGBoost to capture complex non-linear relationships
and interactions that linear models might miss.

### Cross-Validation and Training

We optimize for **AUC-PR** (area under the precision-recall curve),
which directly measures precision vs. recall trade-offs.

```{r}
param_grid <- expand.grid(
  eta = 0.05,
  max_depth = 3,
  subsample = 0.8,
  colsample_bytree = 0.8,
  eval_metric = "aucpr",         # maximize precision-recall AUC
  objective = "binary:logistic"  # output probabilities
)

watchlist = list(train = dtrain)

model <- xgb.train(
  params = as.list(param_grid),
  data = dtrain,
  nrounds = 100,
  watchlist = watchlist,
  verbose = 0
)

importance_matrix <- xgb.importance(model = model)
xgb.plot.importance(importance_matrix[1:10,])
```

------------------------------------------------------------------------

### Threshold Selection Based on Precision

Rather than defaulting to a 0.5 decision threshold, we **optimize
threshold** $T$ to maximize:

$$
\text{Precision} = \frac{TP}{TP + FP}
$$

We assume model predictions $\hat{y} \sim N(\mu, \sigma^2)$ and scan
high quantiles (e.g., 0.99999) to reduce false positives.

```{r}
pred_probs <- predict(model, dtrain)
probs <- 10^seq(-2, -10, length.out = 9)

fit_mean = mean(pred_probs)
fit_sd = sd(pred_probs)

best_precision = 0
best_threshold_xg = 0.5

for (prob in probs){
  threshold <- qnorm(1 - prob, mean = fit_mean, sd = fit_sd)
  preds = ifelse(pred_probs > threshold, 1, 0)
  TP = sum(preds == 1 & train$PL == 1)
  FP = sum(preds == 1 & train$PL == 0)
  if ((TP + FP) < 30) {
    break # CLT
  }
  precision = ifelse((TP + FP) == 0, NA, TP / (TP + FP))
  if (!is.na(precision) && precision > best_precision){
    best_precision = precision
    best_threshold_xg = threshold
  }
  cat(sprintf("Prob = %e | TP = %d | FP = %d | Winrate = %.3f\n", 
              prob, TP, FP, precision))
}

cat("Best Precision:", round(best_precision * 100, 2), "% at Threshold", best_threshold_xg, "\n")
```

### Backtesting Equity Curve

Using the **test set**, we execute trades where the predicted
probability $\hat{y} > T$, and track capital changes.

Let: - $C_0$: Initial capital - $r_i$: Profit/loss from trade $i$ - $T$:
Threshold for high-precision entry

Then: $$
C_{i+1} = C_i + r_i \quad \text{only if } \hat{y}_i > T
$$

```{r}
test_X = model.matrix(PL ~ (ATR + SMA_k7 + SMA_k50 + norm_volume)^4, data=test)[, -1]
dtest = xgb.DMatrix(data = test_X)

pred_probs_xg <- predict(model, dtest)

initial_capital = (1/(0.02)) * (1000 * sample(train$Open, 1))
capital = initial_capital
drawdown = c(capital)
wins = c()

for (i in 1:nrow(test)) {
  pred_X = pred_probs_xg[i]
  if (pred_X > best_threshold_xg) {
    capital = capital + test$pl_value[i]
    drawdown = c(drawdown, capital)
    wins = c(wins, test$pl_value[i])
  }
}

ggplot(data = data.frame(trade = 1:length(drawdown), 
                        equity = drawdown), 
       aes(x = trade, y = equity)) +
  geom_line(color = "blue", size = 1) +
  theme_minimal() +
  labs(title = "XGBoost Equity Curve",
       x = "Trade #",
       y = "Capital") +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey95")
  ) +
  scale_y_continuous(labels = scales::comma)


# Performance metrics
sharpe = (tail(drawdown, 1) - drawdown[1]) / sd(wins)
max_dd = min(drawdown - cummax(drawdown)) / max(drawdown)

cat("Final Capital:", capital, "\n")
cat("Sharpe Ratio:", round(sharpe, 3), "\n")
cat("Max Drawdown:", round(max_dd * 100, 2), "%\n")
```

### Commentary

XGBoost is particularly effective when: - Feature interactions matter -
There are non-linear decision boundaries - **precision over recall**

Its ability to output probabilities makes it well-suited for **threshold
tuning**, which we exploit using quantiles of the fitted distribution.
By focusing on **extremely confident predictions**, we reduce bad trades
and focus capital on high-reward setups.

## Support Vector Machines

Support Vector Machines (SVM) offer a powerful alternative approach for
classification tasks in trading. Unlike linear models, SVMs can
efficiently handle non-linear decision boundaries by projecting data
into higher-dimensional spaces where separation becomes possible.

### Theoretical Foundation

The SVM algorithm finds the optimal hyperplane that maximizes the margin
between classes. For linearly separable data, this involves solving:

$$\min_{\mathbf{w}, b} \frac{1}{2} ||\mathbf{w}||^2$$
$$\text{subject to } y_i(\mathbf{w} \cdot \mathbf{x}_i + b) \geq 1 \text{ for all } i$$

For non-linear boundaries, SVMs employ the "kernel trick" to implicitly
map data to higher dimensions without explicitly computing the
transformation. The decision function becomes:

$$f(\mathbf{x}) = \text{sign}\left(\sum_{i=1}^n \alpha_i y_i K(\mathbf{x}_i, \mathbf{x}) + b\right)$$

Where $K(\mathbf{x}_i, \mathbf{x})$ is the kernel function measuring
similarity between points.

### Model Implementation

RBF (Radial Basis Function) kernel in your SVM is a solid choice for
optimizing precision

```{r}
library(e1071)

X = model.matrix(PL ~ (ATR + SMA_k7 + SMA_k50 + norm_volume)^4, data=train)[, -1]
y = as.factor(train$PL)

svm_model <- svm(x = X, y = y, 
                 kernel = "radial", # RBF is renowned for precision
                 probability = TRUE,
                 cost = 10,
                 gamma = 0.1)

summary(svm_model)
```

### Kernel Selection Rationale

The radial basis function (RBF) kernel is chosen for its ability to
model complex non-linear decision boundaries. The RBF kernel computes:

$$K(\mathbf{x}_i, \mathbf{x}_j) = \exp(-\gamma ||\mathbf{x}_i - \mathbf{x}_j||^2)$$

Where $\gamma$ controls the influence radius of each support vector.
This allows the model to capture local patterns in the feature space
that might be missed by linear models.

### Hyperparameter Tuning
Far too computationally expensive to run, `tune.svm` will take forever to compile. This is an area of improvement.

### Threshold Optimization for Precision

```{r}
svm_probs <- attr(predict(svm_model, X, probability = TRUE), "probabilities")[, "1"]

# Define threshold range
probs <- 10^seq(-2, -10, length.out = 9)
fit_mean <- mean(svm_probs)
fit_sd <- sd(svm_probs)

best_precision <- 0
best_threshold_svm <- 0.5

for (prob in probs) {
  threshold <- qnorm(1 - prob, mean = fit_mean, sd = fit_sd)
  preds <- ifelse(svm_probs > threshold, 1, 0)
  
  TP <- sum(preds == 1 & train$PL == 1)
  FP <- sum(preds == 1 & train$PL == 0)
  
  if ((TP + FP) < 30) {
    break # Ensure statistical significance
  }
  
  precision <- ifelse((TP + FP) == 0, NA, TP / (TP + FP))
  
  if (!is.na(precision) && precision > best_precision) {
    best_precision <- precision
    best_threshold_svm <- threshold
  }
  cat(sprintf("Prob = %e | TP = %d | FP = %d | Winrate = %.3f\n", 
              prob, TP, FP, precision))
}

cat("SVM Best Precision:", round(best_precision * 100, 2), "% at Threshold", best_threshold_svm, "\n")
```

### Backtesting SVM Strategy

```{r}
#test data
test_X <- model.matrix(PL ~ (ATR + SMA_k7 + SMA_k50 + norm_volume)^4, data=test)[, -1]

# Get probability predictions for test set
test_probs_svm <- attr(predict(svm_model, test_X, probability = TRUE), "probabilities")[, "1"]

initial_capital <- (1/(0.02)) * (1000 * sample(train$Open, 1))
capital <- initial_capital
drawdown <- c(capital)
wins <- c()

for (i in 1:nrow(test)) {
  pred_X <- test_probs_svm[i]
  if (pred_X > best_threshold_svm) {
    capital <- capital + test$pl_value[i]
    drawdown <- c(drawdown, capital)
    wins <- c(wins, test$pl_value[i])
  }
}

ggplot(data = data.frame(trade = 1:length(drawdown), 
                        equity = drawdown), 
       aes(x = trade, y = equity)) +
  geom_line(color = "darkgreen", size = 1) +
  theme_minimal() +
  labs(title = "SVM Equity Curve",
       x = "Trade #",
       y = "Capital") +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey95")
  ) +
  scale_y_continuous(labels = scales::comma)

# Calculate performance metrics
sharpe <- (tail(drawdown, 1) - drawdown[1]) / sd(wins)
max_dd <- min(drawdown - cummax(drawdown)) / max(drawdown)

cat("Initial Capital:", initial_capital, "\n")
cat("Final Capital:", capital, "\n")
cat("Sharpe Ratio:", round(sharpe, 3), "\n")
cat("Max Drawdown:", round(max_dd * 100, 2), "%\n")
cat("Returns:", round(((capital/initial_capital)-1)*100, 2), "%\n")
```

### SVM Advantages for Trading

SVMs offer several advantages for financial prediction:

1.  **Robustness to Outliers**: The support vector focus makes SVMs less
    sensitive to outliers compared to regression models.

2.  **Effective in High-Dimensional Spaces**: SVMs perform well with
    many features relative to sample size, ideal for complex market
    data.

3.  **Flexibility**: Through kernel selection, SVMs can model various
    decision boundaries without explicit feature engineering.

4.  **Theoretical Guarantees**: SVMs are founded on statistical learning
    theory, providing formal bounds on generalization error.

The non-linear mapping capability is particularly valuable for capturing
market regimes where relationships between indicators and outcomes
change dynamically.

# Random chance
It's possible that the data is looking at a biased subset of information. It's possible that the overall trend from when this data was collected was upward, and these models wouldn't hold up in turbulent market. To disprove this, A random trader was utilized, trading on a coin toss
```{r}
initial_capital <- (1/(0.02)) * (1000 * sample(train$Open, 1))
capital <- initial_capital
drawdown <- c(capital)
wins <- c()

for (i in 1:nrow(test)) {
  if (round(runif(1,0,1)) == 1) {
    capital <- capital + test$pl_value[i]
    drawdown <- c(drawdown, capital)
    wins <- c(wins, test$pl_value[i])
  }
}

ggplot(data = data.frame(trade = 1:length(drawdown), 
                        equity = drawdown), 
       aes(x = trade, y = equity)) +
  geom_line(color = "black", size = 1) +
  theme_minimal() +
  labs(title = "Monkey trader Equity Curve",
       x = "Trade #",
       y = "Capital") +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey95")
  ) +
  scale_y_continuous(labels = scales::comma)

# Calculate performance metrics
sharpe <- (tail(drawdown, 1) - drawdown[1]) / sd(wins)
max_dd <- min(drawdown - cummax(drawdown)) / max(drawdown)

cat("Initial Capital:", initial_capital, "\n")
cat("Final Capital:", capital, "\n")
cat("Sharpe Ratio:", round(sharpe, 3), "\n")
cat("Max Drawdown:", round(max_dd * 100, 2), "%\n")
cat("Returns:", round(((capital/initial_capital)-1)*100, 2), "%\n")
```

This makes it clear that our models perform better than random chance

# Ensemble prediction
For production trading systems, an ensemble approach combining these
models might offer the most robust performance across varying market
conditions. Each model captures different aspects of market behavior,
and their combined signals could provide more reliable trading decisions
than any single model alone.
```{r}
initial_capital = (1/(0.02))*(1000 * sample(train$Open, 1)) # random price
capital <- initial_capital
drawdown <- c(capital)
wins <- c()

for (i in 1:nrow(test)) {
  pred_X <- predict(glm2, test[i, ], type = "response")
  pred_xg = pred_probs_xg[i]
  pred_svm = test_probs_svm[i]
  # Best threshold not used because all 3 never aligned in dataset
  if (pred_X > 0.5 & pred_xg > 0.5 & pred_svm > 0.5) {
    capital <- capital + test$pl_value[i]
    drawdown <- c(drawdown, capital)
    wins <- c(wins, test$pl_value[i])
  }
}

ggplot(data = data.frame(trade = 1:length(drawdown), 
                        equity = drawdown), 
       aes(x = trade, y = equity)) +
  geom_line(color = "maroon", size = 1) +
  theme_minimal() +
  labs(title = "Equity Curve",
       x = "Trade #",
       y = "Capital") +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey95")
  ) +
  scale_y_continuous(labels = scales::comma)

# Sharpe ratio (simplified version: mean return over SD of returns)
sharpe <- (tail(drawdown, 1) - initial_capital) / sd(wins)


peak = max(drawdown)
trough = min(drawdown)
max_drawdown = (trough-peak)/peak


cat("Initial Capital", initial_capital)
cat("\nFinal Capital:", tail(drawdown, 1), "\n")
cat("Sharpe Ratio:", round(sharpe, 3), "\n")
cat("Max Drawdown:", max_drawdown*100, "%\n")
cat("Returns:", ((capital/initial_capital)-1)*100, "%\n")
```
```

------------------------------------------------------------------------

# Final Thoughts

We have now explored **three models** in depth:

| Model | Output Type | Thresholding | Precision Tuning | Purpose |
|--------------|--------------|--------------|-----------------|--------------|
| Logistic Regression (GLM) | Probabilities (binary) | Yes | qnorm thresholds | Predict trade direction |
| XGBoost | Probabilities (binary) | Yes | precision-optimized threshold | Precision-focused execution |
| Support Vector Machines | Probabilities (binary) | Yes | precision-optimized threshold | Non-linear boundary detection |

Each model offers distinct advantages for trading strategy development:

**GLM** provides interpretable coefficients that directly quantify the
impact of each feature on trade outcomes. This interpretability is
valuable for understanding market dynamics and regulatory compliance.

**XGBoost** excels at capturing complex feature interactions and
non-linear patterns without requiring explicit specification. Its
ensemble nature provides robustness against overfitting, particularly
important in noisy financial data.

**SVM** offers strong theoretical guarantees and effectively handles
non-linear decision boundaries through kernel methods. Its focus on
maximizing the margin between classes helps identify the most reliable
trading signals.

In trading, minimizing **false positives** is often more critical than
maximizing true positives due to the **asymmetric cost of bad trades**.
All three models allow for threshold optimization to achieve
precision-focused execution. These models were put into production, and we found that glm models performed the best in real-time trading.
